---
title: "CPUE Data Wrangle"
author: "Fish Team!"
date: "7/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(ggplot2)
```

##Import Data
```{r import data and load packages, echo=TRUE, include=TRUE, cache=TRUE}
setwd("~/oss/Synthesis/Seamap")
invrec <-read.table("INVREC.txt", header=TRUE,sep=",")
starec <-read.table("STAREC.txt", header=TRUE,sep=",")
bgsrec <-read.table("BGSREC.txt", header=TRUE,sep=",")
glfrec<- read.table("GLFREC.txt", header=TRUE,sep=",")

head(invrec)
head(starec)
head(bgsrec)
head(glfrec)
```

```{r clean and parse down tables, include=FALSE, cache=TRUE}
toNumerics <- function(Date) {
  stopifnot(inherits(Date, c("Date", "POSIXt")))
  day <- as.numeric(strftime(Date, format = "%d"))
  month <- as.numeric(strftime(Date, format = "%m"))
  year <- as.numeric(strftime(Date, format = "%Y"))
  list(year = year, month = month, day = day)
}

fishdata <- dplyr::select(bgsrec,STATIONID,GENUS_BGS,SPEC_BGS,CNTEXP)
starec$Date <- as.POSIXct(starec$MO_DAY_YR,format='%m/%d/%Y')
starec$Time <- sapply(strsplit(as.character(starec$MO_DAY_YR), " "), "[", 2)
testdates<-data.frame(toNumerics(starec$Date))
starec$year<-testdates$year
starec$month<-testdates$month
starec$day<-testdates$day
location <- dplyr::select(starec,STATIONID,DECSLAT,DECSLON,HAULVALUE,year,month,day)
gear <- dplyr::select(invrec,STATIONID, GEAR_SIZE, GEAR_TYPE, MESH_SIZE, MIN_FISH)
rm(bgsrec,invrec,starec,testdates)
size<- glfrec %>% select(STATIONID, GLFID, SPEC_GLF, LEN_GLF) %>% filter(SPEC_GLF=="CAMPEC") #Remove some columns and filter for red snapper
```

##Clean Data

####Fish Data
```{r Display Fish data, echo=TRUE, include=TRUE}
head(fishdata)
```

####Location Data
```{r Display Location data, echo=TRUE, include=TRUE}
head(location)
```

####Gear Data
```{r Display Gear data, echo=TRUE, include=TRUE}
head(gear)
```

####Size Data
```{r Display SIze data, echo=TRUE, include=TRUE}
head(size)
```

##Select Gear Type

####Number of different gear combinations
```{r echo=FALSE}
rs<-fishdata %>% dplyr::filter(SPEC_BGS=="CAMPEC")
rs_gear<- dplyr::inner_join(rs, gear, by="STATIONID")
rs_gear<- rs_gear %>% unite("GEAR_COMBO", c("GEAR_TYPE", "GEAR_SIZE", "MESH_SIZE"))
length(unique(rs_gear$GEAR_COMBO))

barplot(rev(sort(table(rs_gear$GEAR_COMBO))), xlab = "Gear", ylab = "Frequency", main = "Frequency of Gear Used to Catch Red Snapper")
```

```{r Select Gear Type, echo=TRUE, include=TRUE}
gearshrimptrawl <- dplyr::filter(gear, GEAR_TYPE=="ST", GEAR_SIZE==40, MESH_SIZE==1.63) 
head(gearshrimptrawl)
```

####Are We Catching Adults or Juveniles?
```{r Length Frequency}

length_freq<- left_join(select(size, STATIONID, LEN_GLF), select(gear, STATIONID, GEAR_TYPE, GEAR_SIZE, MESH_SIZE), by="STATIONID")

#Check freqency of each gear type
length_freq<- length_freq %>% dplyr::filter(GEAR_TYPE=="ST", GEAR_SIZE==40, MESH_SIZE==1.63, !is.na(LEN_GLF)) 

Lm= 230 #Length at maturity Red Snapper
        #Maturity obtained at year 2, but estimate for size at year 2 is greater than 230cm
```

```{r echo=FALSE, message= FALSE}
#Frequency distribution of Red Snapper

ggplot(length_freq, aes(length_freq$LEN_GLF))+
  geom_histogram()+
  annotate(geom="text",x=235, y=20000, label="Lm", hjust=0)+
  geom_vline(xintercept=230)+
  labs(x="Length (mm)", y="counts", title="Length Distribution")+
  theme(plot.title = element_text(hjust = 0.5))
```

#####Percent Juveniles

```{r echo=TRUE, include=TRUE}
sum(length_freq$LEN_GLF<230)/length(length_freq$LEN_GLF)*100
```

##Join Gear and Location

###Gear
```{r echo=TRUE, include=TRUE}
head(gearshrimptrawl)
```

####Location
```{r echo=TRUE, include=TRUE}
head(location)
```

####Gear + Location
```{r echo=TRUE, include=TRUE}
location_gearST <- dplyr::inner_join(location,gearshrimptrawl,by="STATIONID")
head(location_gearST)
```

##Join Fish Data

####Gear + Location 
```{r echo=TRUE, include=TRUE}
location_gearST <- dplyr::inner_join(location,gearshrimptrawl,by="STATIONID")
head(location_gearST)
```

####Gear + Location + Fish
```{r echo=TRUE, include=TRUE}
location_gearST_fish <- dplyr::left_join(location_gearST,fishdata,by="STATIONID")
head(location_gearST_fish)
```

##Get sites where Red Snapper data was present

####Filter Fish Data for Red Snapper
```{r echo=TRUE, include=TRUE}
redsnapper <- dplyr::filter(fishdata,GENUS_BGS=="LUTJANU",SPEC_BGS=="CAMPEC")
head(redsnapper)
```

##Join Red Snapper

####Gear + Location + Fish +Red Snapper
```{r echo=TRUE, include=TRUE}
location_gearST_redsnapper <- dplyr::left_join(location_gearST,redsnapper,by="STATIONID")
head(location_gearST_redsnapper)
```

##Why didn't we filter for Red Snapper earlier in Fish Data?
#### To detect true absence
#####NA's represent areas where shrimp trawl was used but red snapper not caught
```{r echo=TRUE, include=TRUE}
head(location_gearST_redsnapper)
```
####How many absences?

```{r echo=TRUE, include=TRUE}
sum(is.na(location_gearST_redsnapper$SPEC_BGS))
```

####Add zeros and genus/species
```{r echo=TRUE, include=TRUE}
location_gearST_redsnapper$GENUS_BGS <-"LUTJANU"
location_gearST_redsnapper$SPEC_BGS <-"CAMPEC"
location_gearST_redsnapper[is.na(location_gearST_redsnapper$CNTEXP),"CNTEXP"] <-0
head(location_gearST_redsnapper)
```

##Remove Bad Hauls
```{r echo=TRUE, include=TRUE}
location_gearST_redsnapper <- dplyr::filter(location_gearST_redsnapper,HAULVALUE!="B")
summary(location_gearST_redsnapper$HAULVALUE)
```

##Check for and remove missing data
```{r echo=TRUE, include=TRUE}
colSums(is.na(location_gearST_redsnapper))
location_gearST_redsnapper<- na.omit(location_gearST_redsnapper)
```

##Calculate Catch Per Unit Effort (CPUE)
####CPUE=Count/Minutes Trawled
```{r echo=TRUE, include=TRUE}
location_gearST_redsnapper<- location_gearST_redsnapper %>% dplyr::mutate(CPUE=CNTEXP/MIN_FISH)
head(location_gearST_redsnapper)
```

##Which months were the Red Snapper collected the most?

```{r echo=TRUE, include=TRUE}
counts<-table(location_gearST_redsnapper$month)
barplot(counts, xlab="Month", ylab = "Sampling Effort", main="Red Snapper Sampling Effort vs Month")
```

####June-July and October-November

##Select Data in those windows

```{r warning=FALSE}
location_gearST_redsnapper_sumfall<- location_gearST_redsnapper %>% dplyr::filter(month==6:11)
head(location_gearST_redsnapper_sumfall)
```

##Select Summer Data
```{r echo=TRUE, include=TRUE, warning=FALSE}
location_gearST_redsnapper_summer<- location_gearST_redsnapper %>% dplyr::filter(month==c(6,7))
head(location_gearST_redsnapper_summer)
```

##Summer CPUE Over Time

<iframe width="560" height="315" src="https://nceas.github.io/oss-fishteam/SeamapCPUEdeMutsert.mp4" frameborder="0" allowfullscreen></iframe>

